% =========================================================================
% CHAPTER 2: LITERATURE REVIEW
% =========================================================================
\section{Literature Review}
\label{sec:literature_review}

This chapter surveys the mathematical and computational foundations on
which the thesis rests.  The exposition follows the logical structure of
the hedging problem itself: from the classical theory of replication in
complete markets, through the challenges introduced by market
incompleteness, to the representation theorems that characterise the
minimal dominating wealth process, and finally to the deep learning
architectures that provide computable approximations.  Each section
introduces the economic motivation before the formal mathematics, and
distinguishes clearly between existence results --- theorems guaranteeing
that certain objects exist --- and constructive procedures --- algorithms
that compute or approximate them.

The literature reviewed here spans several fields: mathematical finance,
stochastic analysis, risk management, and machine learning.  In each
case, the focus is on the specific results and methods that are directly
relevant to the problem formulated in Chapter~1: approximating the
predictable integrand~$\tilde{\varphi}$ and the adapted cost
process~$\tilde{C}$ in the optional decomposition of the minimal
dominating wealth process for American options under non-unique martingale
measures.


% =========================================================================
% 2.1  COMPLETE MARKETS AND REPLICATION
% =========================================================================
\subsection{Complete Markets and Replication}
\label{subsec:complete_markets}

The starting point for the theory of derivative pricing is the observation
that, in certain idealised markets, the price of every contingent claim is
determined not by the preferences of individual agents but by the
requirement that no risk-free profit be available.  This section develops
the classical theory of replication in complete markets, establishing the
results that the remainder of the thesis will systematically relax.

The theory is presented in six stages.  First, the economic logic of
arbitrage pricing is explained without heavy notation.  Second, the
probabilistic framework of semimartingales and equivalent measures is laid
out.  Third, the fundamental theorems of asset pricing connect
no-arbitrage to the existence and uniqueness of pricing measures.  Fourth,
the martingale representation theorem is developed as the mathematical
mechanism by which measure uniqueness translates into the existence of
replicating strategies --- this is the step that ensures the cost
process~$\tilde{C}$ vanishes identically.  Fifth, the replication of both
European and American claims is treated, with particular attention to the
Snell envelope and the Doob--Meyer decomposition.  Sixth, the structural
reasons why the replication framework breaks down in practice are
catalogued, motivating the transition to incomplete markets in
Section~\ref{subsec:incomplete_markets}.


% -----------------------------------------------------------------
\subsubsection{Economic Motivation and Arbitrage Pricing}
\label{sssec:arbitrage_pricing}

Consider a financial market in which a derivative security --- say a
European call option --- is available for trade alongside the underlying
asset and a risk-free bond.  The derivative pays a known function of the
asset price at maturity, and the question is: what is its fair price today?

One approach is to model investor preferences, estimate the probability
distribution of future asset prices, and compute the expected discounted
payoff that a particular investor would be willing to accept.  This
approach, rooted in equilibrium theory, requires specifying a utility
function, a subjective probability measure, and an equilibrium condition.
Different investors with different preferences would generally arrive at
different prices, and the resulting theory would be as varied as the
population of market participants.

The insight of the arbitrage pricing approach, developed by
\citet{BlackScholes1973} and \citet{Merton1973}, is that under certain
conditions none of this is necessary.  If the derivative's payoff can be
exactly replicated by a dynamically managed portfolio of the underlying
asset and the bond, then the derivative's price must equal the cost of the
replicating portfolio.  The argument is purely logical.  If the derivative
were priced above the replicating portfolio, one could sell the derivative,
buy the replicating portfolio, and pocket the difference at time zero; at
maturity, the portfolio's payoff would exactly cover the derivative
obligation, leaving a risk-free profit.  If the derivative were priced
below the replicating portfolio, the reverse trade would yield the same
result.  In either case, a risk-free profit emerges from zero initial
investment --- an arbitrage.  Ruling out arbitrage forces the derivative
price to equal the replication cost.

Three conditions are needed for this argument to work.  First, the market
must admit continuous trading in the underlying asset and the bond, so
that the replicating portfolio can be adjusted without restriction.
Second, there must be no transaction costs, so that the adjustment is
costless.  Third, the replicating portfolio must exist: the dynamics of
the underlying asset must be sufficiently rich that the derivative's
payoff can be exactly matched by a trading strategy.

The third condition is the one that defines market completeness.  A
complete market is one in which every contingent claim can be replicated
by trading in the available instruments.  In such a market, every
derivative has a unique arbitrage-free price, and a trading strategy that
produces the payoff with certainty can be explicitly constructed.  The
price is independent of investor preferences because the replication
argument does not involve preferences: it is a logical consequence of the
absence of arbitrage.

The mathematical challenge is to formalise these ideas --- to define
precisely what ``replication'' and ``arbitrage'' mean in continuous time,
and to characterise the class of markets in which completeness holds.
This requires the probabilistic framework developed in the next
subsection.


% -----------------------------------------------------------------
\subsubsection{Mathematical Market Framework}
\label{sssec:prob_framework}

The formal setting is a filtered probability space
$(\Omega, \mathcal{F},
\mathbb{F} := (\mathcal{F}_t)_{t \in [0,T]}, \mathbb{P})$
satisfying the usual conditions of right-continuity and completeness, as
specified in Section~\ref{subsec:math_formulation}.  The sample
space~$\Omega$ represents the set of all possible market scenarios; each
$\omega \in \Omega$ is a complete description of how every price, every
interest rate, and every other market quantity evolves over the trading
horizon~$[0,T]$.  The $\sigma$-algebra~$\mathcal{F}$ is the collection of
all events about which questions can, in principle, be asked.  The
filtration~$\mathbb{F}$ models the flow of information: the
sub-$\sigma$-algebra~$\mathcal{F}_t$ contains every event that can be
determined from observations up to and including time~$t$.
The physical measure~$\mathbb{P}$ assigns probabilities to events and
governs the statistical behaviour of prices as they would be observed in
the real world: it encodes drift rates, volatility levels, correlation
structures, and the frequencies of various market outcomes.

Consider a market consisting of $d+1$ assets: a risk-free money market
account $B_t = e^{rt}$ with constant interest rate $r \ge 0$, and $d$
risky assets with undiscounted price processes $S^1, \ldots, S^d$.  The
money market account serves as the num\'eraire: it defines the unit in
which all other values are expressed.  The discounted price of asset~$i$
is $X^i_t := e^{-rt} S^i_t = S^i_t / B_t$, and the vector
$X := (X^1, \ldots, X^d)$ is assumed to be a locally bounded,
c\`{a}dl\`{a}g semimartingale under~$\mathbb{P}$.  C\`{a}dl\`{a}g means
right-continuous with left limits: at each time, the path is continuous
from the right, and the left-hand limit exists.  This allows for price
jumps (left-hand limit differs from the current value) while ensuring
that the path is well-behaved enough for stochastic calculus.

The semimartingale assumption is the minimal regularity condition under
which a theory of stochastic integration can be developed.  A
semimartingale is a process that admits the canonical decomposition
\begin{equation}
\label{eq:semimartingale_decomp}
X_t \;=\; X_0 \;+\; M_t \;+\; A_t,
\end{equation}
where $M$ is a local martingale (a process with no systematic drift, when
viewed locally) and $A$ is a process of locally finite variation (a
process whose paths have finite total variation on every compact interval,
almost surely).  The decomposition separates the unpredictable
fluctuations of the price ($M$) from its systematic tendency ($A$).
Economically, the local martingale component captures the ``noise'' in
price movements --- the part that cannot be predicted from past
information --- while the finite variation component captures the
``signal'' --- the part that reflects expected returns, risk premia, or
other systematic factors.

The importance of the semimartingale assumption extends beyond technical
convenience.  \citet{DelbaenSchachermayer1994} proved that the
semimartingale property is implied by the absence of arbitrage: if
discounted prices are not semimartingales, then arbitrage opportunities
necessarily exist.  In other words, the semimartingale framework is not an
arbitrary mathematical choice --- it is the largest class of processes
that is compatible with economic equilibrium.

A trading strategy is a predictable, $X$-integrable process
$\varphi := (\varphi^1, \ldots, \varphi^d)$ specifying the number of
units of each risky asset held at each time.  Predictability means that
$\varphi_t$ is $\mathcal{F}_{t^-}$-measurable: the portfolio position at
time~$t$ is determined using only information available strictly
before~$t$.  This is not a mathematical convenience but an economic
necessity --- the hedger must decide how many shares to hold before
observing the next price movement.  Any strategy that could ``see'' the
current price increment before committing to a position would violate the
information structure of the market and correspond to insider trading
rather than legitimate hedging.

Given an initial capital $v \in \mathbb{R}$ and a self-financing
strategy~$\varphi$, the gains-from-trade process is the stochastic
integral
\begin{equation}
\label{eq:gains_from_trade}
G_t(\varphi)
\;:=\; \int_0^t \varphi_s \cdot dX_s
\;=\; \sum_{i=1}^{d} \int_0^t \varphi^i_s \, dX^i_s,
\end{equation}
and the wealth process is $V_t = v + G_t(\varphi)$.  The stochastic
integral $\int_0^t \varphi_s \cdot dX_s$ formalises the cumulative
profit or loss from trading: it accumulates the products of position sizes
and price increments over the trading horizon, in the same way that the
profit from holding $\varphi$ shares over a small interval $[s, s+ds]$ is
approximately $\varphi_s \, dX_s$.  The self-financing condition means
that no external funds are injected or withdrawn: all changes in wealth
are driven by trading gains alone.

The notion of admissibility controls which strategies are permitted.  Two
definitions appear in the literature, and the distinction is relevant
because it affects which representation theorems apply.
\citet{DelbaenSchachermayer1994} define a strategy as admissible if
$V_t \ge -a$ for all $t \in [0,T]$ and some constant $a \ge 0$: the
wealth process is bounded below, but the bound need not be zero.  This
definition is natural for the first fundamental theorem, where the goal is
to characterise the absence of arbitrage.  \citet{Kramkov1996} uses the
stronger requirement that $V_t \ge 0$ for all~$t$: the wealth process
must remain non-negative.  This stronger condition is needed for the
optional decomposition theorem
(Section~\ref{subsec:optional_decomposition}), which characterises the
minimal dominating wealth process.  The non-negativity constraint ensures
that the hedger cannot borrow indefinitely against future gains, and it is
the appropriate condition for the super-hedging and domination problems
studied in this thesis.  Throughout this thesis, admissibility is defined
in the sense of \citet{Kramkov1996}: $V_t \ge 0$ for all $t \in [0,T]$.


% -----------------------------------------------------------------
\subsubsection{Fundamental Theorems of Asset Pricing}
\label{sssec:FTAP}

The probabilistic concept that makes arbitrage pricing work is the change
of probability measure.  Under the physical measure~$\mathbb{P}$,
discounted asset prices typically have non-zero expected returns: investors
demand compensation for bearing risk, so on average discounted prices
drift upward.  A price at time~$t$ that reflects the risk premium earned
by the asset does not, in general, equal the expected value of the
discounted future price.  But the replication argument is indifferent to
expected returns --- it depends only on the volatility structure of price
movements, not their average direction.  If one could find a probability
measure under which the drift vanished --- so that discounted prices had
zero expected change --- then the price of any replicable claim would
equal a simple expectation under that measure.

This is the role of the equivalent martingale measure.  It is a
probability measure that reassigns the likelihood of different market
scenarios so that, on average, discounted prices neither rise nor fall.
The change of measure does not alter which events are possible (it is
equivalent to~$\mathbb{P}$, meaning it agrees on which events have
zero probability); it only adjusts how likely each event is.  Under the
new measure, the market becomes ``fair'' in the precise sense that no
trading strategy has a positive expected gain, which is the mathematical
expression of no-arbitrage.

\begin{definition}[Equivalent local martingale measure]
\label{def:ELMM}
A probability measure $\mathbb{Q}$ defined on
$(\Omega, \mathcal{F})$ is an equivalent local martingale
measure (ELMM) for the discounted price process~$X$ if:
\begin{enumerate}
  \item $\mathbb{Q} \sim \mathbb{P}$, meaning $\mathbb{Q}$ and
    $\mathbb{P}$ have the same null sets: for every event $A \in
    \mathcal{F}$, $\mathbb{P}(A) = 0$ if and only if
    $\mathbb{Q}(A) = 0$.
  \item Each component $X^i$, $i = 1, \ldots, d$, is a local martingale
    under~$\mathbb{Q}$: there exists a localising sequence of stopping
    times $(\tau_n)_{n \ge 1}$ with $\tau_n \uparrow \infty$ such that
    the stopped process $X^{i,\tau_n}$ is a $\mathbb{Q}$-martingale for
    each~$n$.
\end{enumerate}
The set of all such measures is denoted~$\mathcal{M}(X)$.
\end{definition}

The equivalence condition $\mathbb{Q} \sim \mathbb{P}$ is economically
substantive.  It ensures that the pricing measure and the physical
measure agree on which events are possible: every market scenario that
could occur in the real world ($\mathbb{P}(A) > 0$) receives positive
probability under~$\mathbb{Q}$, and conversely.  A pricing measure that
assigned zero probability to a scenario that could actually materialise
would be ignoring a real risk; a pricing measure that assigned positive
probability to an impossible scenario would be introducing a phantom
risk.  Equivalence prevents both pathologies.

The local martingale condition is a weakening of the full martingale
condition.  A process is a martingale if its conditional expectation at
any future time equals its current value: $\mathbb{E}^{\mathbb{Q}}
[X_t \mid \mathcal{F}_s] = X_s$ for all $s \le t$.  A local martingale
only requires this property to hold locally, after stopping the process at
times that prevent it from becoming too large.  The distinction matters
for continuous-time models with unbounded prices, where the full
martingale property may fail due to integrability issues (e.g.,
exponential martingales that are only local martingales).  For the locally
bounded processes considered in this thesis, the distinction is primarily
technical rather than economic.  In many diffusion models, discounted
prices are true martingales under suitable integrability conditions; in
general semimartingale settings, local martingale is the correct minimal
requirement.

The relationship between the physical and pricing measures is governed by
the Radon--Nikod\'ym theorem: since $\mathbb{Q} \sim \mathbb{P}$, there
exists a positive random variable $L_T := d\mathbb{Q}/d\mathbb{P}
|_{\mathcal{F}_T}$ such that
$\mathbb{E}^{\mathbb{Q}}[Y] = \mathbb{E}^{\mathbb{P}}[L_T Y]$ for
every bounded $\mathcal{F}_T$-measurable random variable~$Y$.  The
density process $L_t := \mathbb{E}^{\mathbb{P}}[L_T \mid \mathcal{F}_t]$
is a positive $\mathbb{P}$-martingale with $L_0 = 1$.  Intuitively, $L_t$
represents the ``exchange rate'' between the two measures at time~$t$: it
re-weights the likelihood of different scenarios to remove the risk
premium from asset prices.

The two fundamental theorems of asset pricing connect these mathematical
objects to the economic structure of the market.  The development of these
theorems spans three decades and several increasingly general
formulations, each motivated by the limitations of the previous one.  The
progression is worth tracing because it illuminates the precise
mathematical conditions under which the theory applies and identifies
where generalisation is still needed for the models in this thesis.

\citet{HarrisonKreps1979} established the first fundamental connection
between no-arbitrage and the existence of a pricing measure, working in a
discrete-time framework with a finite number of trading periods and states.
In their setting, a simple no-arbitrage condition suffices, and the pricing
measure is a full martingale measure (not merely a local martingale
measure).  \citet{HarrisonPliska1981} extended this framework to allow for
continuous trading in a single period, proved the second fundamental
theorem connecting uniqueness to completeness, and established the
martingale characterisation of attainable claims.  However, both
formulations rely on assumptions --- finite state spaces or specific
process structures --- that are too restrictive for continuous-time
models driven by Brownian motion.

The definitive generalisation was provided by
\citet{DelbaenSchachermayer1994}, who worked in the full generality of
locally bounded semimartingales.  Their contribution was twofold.  First,
they showed that simple no-arbitrage is insufficient in continuous time: a
stronger condition, no free lunch with vanishing risk (NFLVR), is
required.  NFLVR rules out not only exact arbitrage but also sequences of
approximate arbitrages whose risk vanishes in the limit --- a pathology
that cannot occur in discrete time but arises naturally in
continuous-time models.  Second, they showed that the pricing measure need
only make discounted prices into local martingales, not full
martingales, because the full martingale property may fail for unbounded
processes even when no arbitrage is present.  These two generalisations
--- from no-arbitrage to NFLVR, and from martingale measures to local
martingale measures --- are what make the theory applicable to the
continuous semimartingale models used in this thesis.

The progression from \citet{HarrisonKreps1979} through
\citet{HarrisonPliska1981} to \citet{DelbaenSchachermayer1994} thus
represents not merely a chronological refinement but a systematic
weakening of hypotheses---from finite state spaces to general
semimartingales, from exact martingale measures to local martingale
measures, and from simple no-arbitrage to NFLVR---each step expanding
the class of models to which the fundamental theorems apply while
preserving their economic content.  The cost of this generality is that
constructive tools (explicit Girsanov transforms, closed-form densities)
are available only in specific parametric models, a limitation that
becomes acute in the incomplete-market setting of
Section~\ref{subsec:incomplete_markets}.

\begin{theorem}[First fundamental theorem of asset pricing
{\citep{DelbaenSchachermayer1994}}]
\label{thm:FTAP1}
Let $X$ be a locally bounded, c\`{a}dl\`{a}g semimartingale.  The
market satisfies no free lunch with vanishing risk (NFLVR) if and only
if $\mathcal{M}(X) \neq \emptyset$.
\end{theorem}

\noindent What the theorem says mathematically.
The theorem establishes a bijective correspondence between an economic
property of the market (NFLVR) and a mathematical property of a set
of probability measures ($\mathcal{M}(X) \neq \emptyset$).  The ``only
if'' direction --- NFLVR implies the existence of at least one ELMM ---
is the mathematically deep part; the proof constructs a measure from the
economic assumption using functional analysis (specifically, a separation
argument in $L^\infty$).  The ``if'' direction is more straightforward:
under any~$\mathbb{Q} \in \mathcal{M}(X)$, the gains-from-trade process
$G_t(\varphi) = \int_0^t \varphi_s \cdot dX_s$ is a local martingale (as
a stochastic integral with respect to a local martingale) and, under the
admissibility condition $V_t \ge 0$, a supermartingale with
$G_0(\varphi) = 0$.  Therefore
$\mathbb{E}^{\mathbb{Q}}[G_T(\varphi)] \le 0$ for every admissible
strategy, which precludes strategies with non-negative, strictly positive
payoffs.

\noindent What the theorem means financially.
The theorem transforms an economic question --- ``can anyone make
risk-free profits in this market?'' --- into a mathematical one ---
``does the set~$\mathcal{M}(X)$ contain at least one element?''  The
existence of a pricing measure certifies the economic coherence of the
model: prices are consistent with the absence of arbitrage.  The measure
itself is a computational device; it is not a physical entity and it need
not represent any agent's beliefs about the world.

\noindent What the theorem does not provide.
The theorem is an existence result: it guarantees that at least one ELMM
exists but does not construct it.  In specific models, the measure can be
identified explicitly through the Girsanov theorem (as we demonstrate
below for the Black--Scholes model).  In general, the theorem provides a
theoretical guarantee but not a recipe.  Moreover, the theorem says
nothing about uniqueness: $\mathcal{M}(X)$ may contain one element or
uncountably many.  The question of uniqueness --- which turns out to be
equivalent to market completeness --- is addressed by the second
fundamental theorem.

\begin{theorem}[Second fundamental theorem of asset pricing
{\citep{HarrisonPliska1981}}]
\label{thm:FTAP2}
Let $X$ be a locally bounded semimartingale satisfying NFLVR.  The
market is complete if and only if $|\mathcal{M}(X)| = 1$.
\end{theorem}

\noindent What the theorem says mathematically.
The theorem provides a characterisation of market completeness in terms
of the cardinality of the set of equivalent local martingale measures:
completeness is equivalent to uniqueness.  The ``only if'' direction uses
a separation argument: if two distinct measures $\mathbb{Q}_1,
\mathbb{Q}_2 \in \mathcal{M}(X)$ existed, one could construct a claim
$H = \mathbf{1}_A$ for an event $A$ with
$\mathbb{Q}_1(A) \neq \mathbb{Q}_2(A)$; this claim would have two
different arbitrage-free prices and hence could not be replicable.  The
``if'' direction is the deeper half and relies on the martingale
representation theorem, developed in the next subsection: when only one
measure exists, the filtration is generated by the price process alone, and
every $\mathbb{Q}$-martingale can be written as a stochastic integral with
respect to~$X$.

\noindent What the theorem means financially.
In a complete market, there is no ambiguity in derivative pricing.  Every
derivative has exactly one arbitrage-free price, and a replicating
strategy exists for every payoff.  The price can be computed as a single
expectation under the unique measure~$\mathbb{Q}$:
\begin{equation}
\label{eq:risk_neutral_pricing}
\pi(H) \;=\; \mathbb{E}^{\mathbb{Q}}[e^{-rT} H],
\end{equation}
and the replicating strategy can be extracted from the martingale
representation of the discounted claim (see
Section~\ref{sssec:mart_rep}).  Every market participant, using the same
model, must arrive at the same price.

When $|\mathcal{M}(X)| > 1$, the market is incomplete: some claims are
not replicable, and different measures in~$\mathcal{M}(X)$ assign them
different prices.  The unique price~\eqref{eq:risk_neutral_pricing}
gives way to the price interval~\eqref{eq:price_interval}, and the
replication strategy gives way to the approximate hedging strategies that
are the subject of this thesis.

\noindent Computational implication.
When the market is complete, the computational problem of pricing and
hedging is fully solved in principle: compute the single
expectation~\eqref{eq:risk_neutral_pricing} and differentiate to
obtain the hedge ratios.  The optional decomposition of
Chapter~1, equation~\eqref{eq:optional_decomp}, reduces to a
self-financing replication: the cost process $\tilde{C} \equiv 0$, and
the entire value process is generated by the gains-from-trade integral.
This degenerate case serves as a baseline for the experiments in
Chapter~5: a correctly functioning neural network hedger should produce
near-zero residual error in the complete-market limit, confirming that
the architecture can learn the predictable integrand when no adapted
cost component is present.

\noindent Structural limitation.
The second fundamental theorem characterises completeness but does not
prescribe how to verify it in a given model.  In practice, one checks
completeness by counting dimensions: if the filtration is generated by a
$d$-dimensional Brownian motion and there are $d$ traded risky assets
whose volatility matrix is non-degenerate, the martingale representation
theorem (next subsection) guarantees that every martingale is a stochastic
integral with respect to the price process, and the market is complete.
When the number of Brownian drivers exceeds the number of traded assets,
or when additional state variables (such as stochastic volatility) are
present, the representation fails and the market is incomplete.  This
dimensional diagnostic is the basis for the two-model design of this
thesis: the multi-factor GBM model creates a dimensional mismatch
directly, while the Heston model introduces an untraded state variable.


% -----------------------------------------------------------------
\subsubsection{Martingale Representation and the Predictable Integrand}
\label{sssec:mart_rep}

The fundamental theorems establish when replication is possible
(when $|\mathcal{M}(X)| = 1$), but they do not explain how the
replicating strategy is constructed.  The mathematical mechanism that
converts measure uniqueness into an explicit trading strategy is the
martingale representation theorem.  This result is the reason the cost
process~$\tilde{C}$ vanishes in complete markets, and its failure in
incomplete markets is the reason~$\tilde{C}$ becomes strictly positive.
Because this mechanism is central to the interpretation of the neural
network output throughout the thesis, it merits a dedicated treatment.

\paragraph{The predictable representation property.}
Let $W = (W^1, \ldots, W^d)$ be a $d$-dimensional Brownian motion on
$(\Omega, \mathcal{F}, \mathbb{Q})$, and let $\mathbb{F}^W$ denote the
(augmented) filtration generated by~$W$.  The predictable representation
property states that every $(\mathbb{F}^W, \mathbb{Q})$-local martingale
$N$ can be written as a stochastic integral with respect to~$W$:
\begin{equation}
\label{eq:PRP}
N_t \;=\; N_0 \;+\; \sum_{j=1}^{d} \int_0^t \psi^j_s \, dW^j_s,
\end{equation}
for some predictable, square-integrable process
$\psi = (\psi^1, \ldots, \psi^d)$.  The key condition is that
$\mathbb{F}^W$ is the filtration generated by~$W$: no additional sources
of information are present beyond the Brownian motion itself.

\noindent What the theorem says mathematically.
Every source of randomness in the model --- every martingale adapted to
the Brownian filtration --- can be decomposed into a weighted sum of
stochastic integrals against the individual Brownian components.  The
integrand~$\psi$ is uniquely determined up to
$\mathbb{Q} \otimes dt$-null sets.  This is not an approximation result;
it is an exact decomposition: nothing is left over.

\noindent Financial interpretation.
If the market filtration is generated by a $d$-dimensional Brownian motion
and there are $d$ traded risky assets whose volatility matrix
$\sigma := (\sigma^{ij})$ is invertible, then any claim that depends on
the Brownian motion can be replicated by trading in the $d$ assets.  The
key step is a change of integrand: from the representation~\eqref{eq:PRP}
in terms of~$W$, one obtains a representation in terms of the price
process~$X$ by the transformation $\varphi = \psi \, \sigma^{-1}$.  The
invertibility of~$\sigma$ is what allows the translation from Brownian
integrals to trading integrals.

\paragraph{Application to the Black--Scholes model.}
Under the physical measure~$\mathbb{P}$, the price of a single risky
asset follows geometric Brownian motion:
\begin{equation}
\label{eq:GBM_P}
dS_t \;=\; \mu \, S_t \, dt \;+\; \sigma \, S_t \, dW_t^{\mathbb{P}},
\end{equation}
where $\mu \in \mathbb{R}$ is the drift (expected return per unit time),
$\sigma > 0$ is the constant volatility, and $W^{\mathbb{P}}$ is a
standard Brownian motion under~$\mathbb{P}$.  The model has one traded
risky asset and one source of randomness: the single Brownian motion
$W^{\mathbb{P}}$.  This one-to-one correspondence between instruments and
risk factors is what makes the model complete.

The Girsanov theorem provides the constructive tool for identifying the
unique equivalent local martingale measure.  Define the market price of
risk
\begin{equation}
\label{eq:market_price_risk_BSM}
\gamma \;:=\; \frac{\mu - r}{\sigma},
\end{equation}
which measures the excess return per unit of volatility --- the
compensation that the market demands for bearing one unit of the asset's
risk.  Define the Radon--Nikod\'ym density
\begin{equation}
\label{eq:girsanov_BSM}
\frac{d\mathbb{Q}}{d\mathbb{P}} \bigg|_{\mathcal{F}_T}
\;=\; \exp\!\Bigl(-\gamma \, W_T^{\mathbb{P}}
      - \tfrac{1}{2} \gamma^2 T\Bigr)
\;=:\; L_T.
\end{equation}
The process $L_t := \mathbb{E}^{\mathbb{P}}[L_T \mid \mathcal{F}_t]
= \exp(-\gamma W_t^{\mathbb{P}} - \frac{1}{2}\gamma^2 t)$ is a positive
$\mathbb{P}$-martingale with $L_0 = 1$, and therefore $\mathbb{Q}$
defined by $d\mathbb{Q} = L_T \, d\mathbb{P}$ is a probability measure
equivalent to~$\mathbb{P}$.  By the Girsanov theorem, the process
\begin{equation}
\label{eq:girsanov_W}
W_t^{\mathbb{Q}} \;:=\; W_t^{\mathbb{P}} + \gamma \, t
\end{equation}
is a standard Brownian motion under~$\mathbb{Q}$.  Substituting
into~\eqref{eq:GBM_P}, the discounted price
$X_t = e^{-rt} S_t$ satisfies
\begin{equation}
\label{eq:GBM_Q}
dX_t \;=\; \sigma \, X_t \, dW_t^{\mathbb{Q}}.
\end{equation}
The drift has been removed: under~$\mathbb{Q}$, the discounted price is a
(geometric) martingale.

The construction is unique because the Girsanov theorem requires a
one-dimensional market price of risk~$\gamma$ to satisfy one equation
($\sigma \gamma = \mu - r$), yielding a unique
solution~\eqref{eq:market_price_risk_BSM}.  This confirms
$|\mathcal{M}(X)| = 1$.  The uniqueness is a direct consequence of the
dimensional balance: one equation, one unknown, one solution.

\paragraph{From representation to replication.}
Given the unique measure~$\mathbb{Q}$, the price of a European claim
with payoff $H = h(S_T)$ is
\begin{equation}
\label{eq:BSM_price}
\pi_t \;=\; e^{-r(T-t)} \,
\mathbb{E}^{\mathbb{Q}}[h(S_T) \mid \mathcal{F}_t]
\;=:\; u(t, S_t),
\end{equation}
where $u : [0,T] \times (0,\infty) \to \mathbb{R}$ is the pricing
function.  The discounted claim price $e^{-rt} u(t, S_t)$ is a
$\mathbb{Q}$-martingale adapted to the filtration of~$W^{\mathbb{Q}}$.
By the predictable representation property~\eqref{eq:PRP}, it admits the
decomposition
\begin{equation}
\label{eq:mart_rep_BSM}
e^{-rt} u(t, S_t) \;=\; u(0, S_0)
\;+\; \int_0^t \psi_s \, dW_s^{\mathbb{Q}}
\end{equation}
for a unique predictable process~$\psi$.  Since
$dX_t = \sigma X_t \, dW_t^{\mathbb{Q}}$, the substitution
$\varphi_t := \psi_t / (\sigma X_t)$ gives the equivalent representation
in terms of the price process:
\begin{equation}
\label{eq:rep_in_X}
e^{-rt} u(t, S_t) \;=\; u(0, S_0)
\;+\; \int_0^t \varphi_s \, dX_s.
\end{equation}
It\^o's formula applied to $e^{-rt} u(t, S_t)$ identifies the integrand
explicitly:
\begin{equation}
\label{eq:delta_hedge}
\varphi_t \;=\; \frac{\partial u}{\partial S}(t, S_t).
\end{equation}
This is the classical delta: the sensitivity of the option price to a
change in the underlying asset price.  The strategy that holds
$\varphi_t$ units of the risky asset at each time~$t$ perfectly
replicates the claim, and the cost of the replicating portfolio equals
the option price $\pi_0 = u(0, S_0)$.

\paragraph{Connection to the optional decomposition.}
Equation~\eqref{eq:rep_in_X} is the complete-market instance of the
optional decomposition~\eqref{eq:optional_decomp}.  Comparing the two
makes the role of each component transparent.  In the optional
decomposition $\tilde{V}_t = \tilde{V}_0 + \int_0^t \tilde{\varphi}_s
\, dX_s - \tilde{C}_t$, the initial capital~$\tilde{V}_0$ corresponds
to~$\pi_0$, the predictable integrand~$\tilde{\varphi}$ corresponds to
the delta~$\varphi$, and the cost process~$\tilde{C}$ is identically
zero.  The vanishing of~$\tilde{C}$ is a direct consequence of the
predictable representation property: because every martingale in the
Brownian filtration can be written as a stochastic integral against~$X$,
nothing is left over that the trading strategy cannot capture.

In the standard Brownian setting with non-degenerate volatility and
filtration generated by the driving noise, the chain of logic is:
\begin{center}
\fbox{\parbox{0.85\textwidth}{
$|\mathcal{M}(X)| = 1$
\;$\Longrightarrow$\;
filtration $= \mathbb{F}^W$ (under modelling assumptions)
\;$\Longrightarrow$\;
predictable representation~\eqref{eq:PRP}
\;$\Longrightarrow$\;
$\tilde{C} \equiv 0$.
}}
\end{center}
When any link in this chain breaks --- when the filtration contains
information not generated by the traded assets' Brownian motions, or when
the volatility matrix is not invertible --- the representation fails, and
$\tilde{C}$ becomes strictly positive.  The multi-factor GBM model breaks
the chain at the second step (the filtration contains extra Brownian
motions); the Heston model breaks it at both the first and second steps
(the filtration contains the variance process, and the set of martingale
measures is non-singleton).

\paragraph{Computational cost of the delta hedge.}
For the European call with payoff $h(S_T) = (S_T - K)^+$, the pricing
function $u$ is the Black--Scholes formula
\begin{equation}
\label{eq:BS_call}
u(t, S) \;=\; S \, \Phi(d_+) \;-\; K e^{-r(T-t)} \Phi(d_-),
\end{equation}
where $\Phi$ is the standard normal CDF,
$d_{\pm} := \frac{\ln(S/K) + (r \pm \sigma^2/2)(T-t)}{\sigma\sqrt{T-t}}$,
and the delta is $\varphi_t = \Phi(d_+(t, S_t))$.  The European put
follows by put--call parity.  In this setting, computing the hedge ratio
at each rebalancing date is trivial: evaluate a closed-form expression
involving the normal CDF.

This computational simplicity is specific to the Black--Scholes model.
Even within the class of complete markets, moving to multi-asset models or
path-dependent payoffs eliminates closed-form expressions for the
delta~\eqref{eq:delta_hedge}.  The pricing function $u$ must then be
obtained by solving a PDE numerically (finite differences, finite
elements) or by Monte Carlo estimation, and the delta must be extracted by
numerical differentiation --- a procedure that is noisy when the
underlying estimator has statistical error and expensive when the state
space is high-dimensional.  In the multi-asset models of this thesis, the
state space is at least four-dimensional (two asset prices and two
variance processes), and PDE methods become impractical due to the curse
of dimensionality \citep{Glasserman2003}.  This computational barrier is
one of the primary motivations for approximating~$\tilde{\varphi}$ with
neural networks rather than solving for it analytically or via grid-based
methods: the neural network learns a mapping from state to portfolio
weight that can be evaluated at negligible cost per path, regardless of
the dimension of the state space.


% -----------------------------------------------------------------
\subsubsection{Replication of European and American Claims}
\label{sssec:replication_claims}

With the martingale representation theorem in hand, the replication of
European claims in complete markets is immediate: the discounted claim
price is a $\mathbb{Q}$-martingale, the representation theorem
decomposes it as a stochastic integral, and the integrand is the hedging
strategy.  American claims require an additional step because the exercise
time is not fixed in advance.

\paragraph{European claims.}
For a European claim $H = h(S_T)$ in the Black--Scholes model, the
replication result is:
\begin{equation}
\label{eq:european_replication}
e^{-rT} H \;=\; \pi_0
\;+\; \int_0^T \varphi_s \, dX_s,
\qquad
\varphi_t = \frac{\partial u}{\partial S}(t, S_t),
\end{equation}
where $\pi_0 = u(0, S_0)$ and $u$ is the pricing function.  The
replication is exact: the strategy $\varphi$ produces the payoff $H$
with probability one, and no residual cost remains.

\paragraph{American claims: the optimal stopping problem.}
Consider an American put option with strike~$K$ and maturity~$T$ on
asset~$i$ in a complete market with unique measure~$\mathbb{Q}$.  The
discounted reward process is
\begin{equation}
\label{eq:american_reward_ch2}
\xi_t \;:=\; e^{-rt}(K - S^i_t)^+, \qquad t \in [0,T].
\end{equation}
The holder's problem is to choose a stopping time $\tau \in
\mathcal{T}_{[0,T]}$ that maximises the expected discounted payoff.
The fair price at time~$t$ is
\begin{equation}
\label{eq:american_price_complete}
\pi_t(\xi) \;=\;
\esssup_{\tau \in \mathcal{T}_{[t,T]}}
\mathbb{E}^{\mathbb{Q}}[\xi_\tau \mid \mathcal{F}_t].
\end{equation}

\paragraph{The Snell envelope.}
The value process $(\pi_t(\xi))_{t \in [0,T]}$ is the Snell
envelope of the reward process~$\xi$ under~$\mathbb{Q}$.  The concept
was introduced by \citet{Snell1952} in the context of sequential decision
theory.  The Snell envelope is the smallest $\mathbb{Q}$-supermartingale
that dominates the reward process: $\pi_t(\xi) \ge \xi_t$ for all~$t$,
and any other $\mathbb{Q}$-supermartingale $Y$ with $Y_t \ge \xi_t$ for
all~$t$ also satisfies $Y_t \ge \pi_t(\xi)$ for all~$t$.

The existence of this smallest dominating supermartingale is not
self-evident.  It relies on the fact that the class of
$\mathbb{Q}$-supermartingales dominating~$\xi$ is closed under a
lattice-type operation: the essential infimum of any family of
supermartingales dominating~$\xi$ is itself a supermartingale
dominating~$\xi$, provided the family is directed downward.  For reward
processes that are right-continuous and of class~(D) --- conditions
satisfied by the discounted American put payoff~\eqref{eq:american_reward_ch2},
which is non-negative and bounded above by~$K$ --- this closure property
holds, and the Snell envelope is well-defined.  A detailed treatment can
be found in \citet{KaratzasShreve1998} and \citet{Peskir2006}.

\noindent Financial interpretation.  The Snell envelope is the
minimum amount of wealth that the option seller must maintain to guarantee
coverage of the American option at every point in time.  At each
instant, $\pi_t(\xi)$ is at least as large as the exercise value~$\xi_t$
(the seller must be able to pay if the holder exercises now) and equals
the conditional expectation of the best future exercise value (if it is
optimal to continue, the seller can afford to do so).

The optimal exercise time is the first time the Snell envelope coincides
with the reward process:
\begin{equation}
\label{eq:optimal_stopping_time}
\tau^* \;:=\; \inf\{t \ge 0 : \pi_t(\xi) = \xi_t\}.
\end{equation}
At $\tau^*$, the option holder exercises because the value of holding is
exactly equal to the value of exercising --- there is no benefit to
waiting further.  The stopped process
$(\pi_{t \wedge \tau^*}(\xi))_{t \in [0,T]}$ is a $\mathbb{Q}$-martingale:
once the stopping decision has been made, the residual randomness is
removed.

\paragraph{Doob--Meyer decomposition and hedging.}
The continuous-time theory of American option pricing was developed by
\citet{Bensoussan1984} and \citet{Karatzas1988}, who established the
connection between the Snell envelope, free-boundary problems, and
variational inequalities.  An important alternative perspective was
provided by the early exercise premium decomposition of
\citet{Kim1990} and \citet{CarrJarrowMyneni1992}, which represents the
American option price as the sum of the corresponding European option
price and an integral over the exercise boundary.  This decomposition
provides economic insight --- the early exercise premium quantifies the
additional value created by the right to exercise before maturity --- and
has been used for analytical approximations.  However, the
Snell-envelope/Doob--Meyer framework is the one that connects directly to
the optional decomposition, and it is therefore the approach adopted in
this thesis.

In a complete market, the Snell envelope admits the Doob--Meyer
decomposition:
\begin{equation}
\label{eq:doob_meyer_american_ch2}
\pi_t(\xi) \;=\; \pi_0(\xi)
\;+\; \int_0^t \varphi_s \, dX_s \;-\; A_t,
\end{equation}
where $\varphi$ is a predictable integrand (the hedging strategy) and
$A$ is a predictable, non-decreasing process with $A_0 = 0$.  The
uniqueness of this decomposition is guaranteed by the Doob--Meyer theorem
\citep{KaratzasShreve1991}: for a c\`{a}dl\`{a}g supermartingale of
class~(D), the decomposition into a martingale and a predictable
increasing process is unique.

The process $A$ increases only on the contact set
$\{t : \pi_t(\xi) = \xi_t\}$, that is, at times when the Snell envelope
meets the payoff process: it reflects the cumulative cost of the early
exercise feature.  Both $\varphi$ and $A$ are predictable: the hedger can
anticipate the trading strategy and the associated cost one instant ahead.

The predictability of $A$ is the key structural feature that distinguishes
the complete-market decomposition from the optional decomposition that
arises in incomplete markets.  In the notation of
equation~\eqref{eq:optional_decomp}, the complete-market
decomposition~\eqref{eq:doob_meyer_american_ch2} corresponds to
$\tilde{V}_t = \pi_t(\xi)$, $\tilde{\varphi}_t = \varphi_t$, and
$\tilde{C}_t = A_t$, with the additional property that $\tilde{C} = A$
is predictable.  In incomplete markets
(Theorem~\ref{thm:optional_decomp} in
Section~\ref{subsec:optional_decomposition}), this predictability is
lost: the cost process becomes merely adapted, reflecting the fact that
some costs cannot be foreseen when unhedgeable risks are present.

\paragraph{Free-boundary problem.}
For the American put in the Black--Scholes model, the pricing function
$u(t, S) = \pi_t(\xi)$ when $\xi_t = e^{-rt}(K - S_t)^+$ satisfies a
free-boundary problem.  In the continuation region
$\mathcal{C} := \{(t,S) : u(t,S) > (K-S)^+\}$, the pricing function
satisfies the Black--Scholes PDE
\begin{equation}
\label{eq:BS_PDE}
\frac{\partial u}{\partial t}
+ \frac{1}{2}\sigma^2 S^2
  \frac{\partial^2 u}{\partial S^2}
+ r S \frac{\partial u}{\partial S}
- r u \;=\; 0.
\end{equation}
In the exercise region
$\mathcal{E} := \{(t,S) : u(t,S) = (K-S)^+\}$, the option is exercised
immediately.  The boundary $S^*(t)$ separating the two regions is not
known in advance and must be determined as part of the solution: it is
the free boundary or optimal exercise boundary.  At the
free boundary, the pricing function satisfies the smooth-pasting
conditions
\begin{equation}
\label{eq:smooth_pasting}
u(t, S^*(t)) = K - S^*(t),
\qquad
\frac{\partial u}{\partial S}(t, S^*(t)) = -1.
\end{equation}

This free-boundary problem does not admit a closed-form solution, and
numerical methods are required.  The standard approaches include finite
difference methods applied to the variational
inequality~\eqref{eq:BS_PDE}, binomial and trinomial tree methods
\citep{CoxRossRubinstein1979}, and Monte Carlo methods with regression.
Finite difference and tree methods discretise the state space on a grid
and solve the PDE or its discrete analogue backward in time.  These
methods are efficient and accurate in low dimensions (one or two state
variables) but scale poorly with the dimension of the state space: the
number of grid points grows exponentially with dimension, rendering the
approach impractical for the four-dimensional state space of the two-asset
Heston model.

\paragraph{Monte Carlo methods: Longstaff--Schwartz.}
The Longstaff--Schwartz algorithm \citep{LongstaffSchwartz2001} is the
standard Monte Carlo approach to American option pricing.  Unlike
grid-based methods, it scales to high-dimensional problems because the
computational cost is linear in the number of simulated paths rather than
exponential in the dimension.  The algorithm works backward in time,
estimating the conditional expectation of the continuation value at each
exercise date by least-squares regression against a set of basis functions
of the current state.

At each exercise date~$t_k$, the algorithm has access to the current
state $S_{t_k}$ and the estimated continuation value
$\hat{C}_{t_{k+1}}$.  The regression
\begin{equation}
\label{eq:LS_regression}
\hat{C}_{t_k}
\;=\; \sum_{j=1}^{J} \beta_j \, \psi_j(S_{t_k})
\end{equation}
projects the future value onto a finite basis
$\{\psi_1, \ldots, \psi_J\}$ (typically polynomials of the state
variables), and the exercise decision is: exercise at $t_k$ if
$\xi_{t_k} \ge \hat{C}_{t_k}$, otherwise continue.  The convergence of
the algorithm as the number of paths and basis functions increase was
established by \citet{TsitsiklisVanRoy2001} and
\citet{ClementLamberton2002}.

The Longstaff--Schwartz algorithm solves the optimal stopping problem but
does not directly produce a hedging strategy.  It outputs the exercise
boundary (or an approximation of it) and the option price, but not the
portfolio weights that would be needed to replicate the option value.
Extracting a hedging strategy from the Longstaff--Schwartz output
requires a separate computation --- typically finite differences of the
estimated continuation value function, using a ``bump and revalue''
procedure --- which introduces an additional layer of approximation
error that compounds with the regression error in the continuation value
estimate.

\paragraph{Comparison of hedging methodologies in complete markets.}
It is worth pausing to compare the three approaches to hedging that have
emerged in this section, since the same structural trade-offs reappear in
amplified form once the market becomes incomplete.

The PDE/analytical approach (delta hedging via
equation~\eqref{eq:delta_hedge}) is exact when a closed-form pricing
function exists and fast to evaluate, but it requires the pricing PDE to
be solved first---either analytically or on a grid---and its
computational cost scales exponentially with the dimension of the state
space \citep{Glasserman2003}.  For the one-dimensional Black--Scholes
model it is the method of choice; for the four-dimensional state space of
the two-asset Heston model studied in Chapter~4, it is infeasible.

The tree/finite-difference approach
\citep{CoxRossRubinstein1979,Brennan1977} discretises the state space
and solves backward in time.  It handles American exercise naturally
through the dynamic-programming recursion and produces both prices and
hedge ratios on the grid.  However, its cost is again exponential in
dimension, and interpolation between grid points introduces additional
error in the hedge ratios.

The Monte Carlo/regression approach (Longstaff--Schwartz) escapes
the curse of dimensionality for pricing---its cost is linear in the
number of paths---but decouples the pricing and hedging steps.  The
hedging strategy must be recovered by finite-difference ``bump and
revalue'' applied to the estimated continuation value, a procedure that
compounds regression error with numerical differentiation noise.
\citet{BroadieGlasserman2004} quantify the bias introduced by the
high/low estimator construction, and \citet{GlassermanYu2004} analyse the
effect of basis-function choice on convergence, but neither addresses the
hedging problem directly.

A common limitation unites all three approaches: they treat pricing as
the primary problem and extract the hedge as a derivative of the
pricing solution---literally, a sensitivity or finite difference.  This
indirect path means that errors in the price estimate propagate into, and
are often amplified by, the hedging computation.  The deep hedging
framework introduced in Section~\ref{subsec:deep_hedging} inverts this
hierarchy, learning the hedging strategy directly from simulated
gain/loss outcomes without requiring a pricing step as an intermediate.


% -----------------------------------------------------------------
\subsubsection{Structural Breakdown of Replication}
\label{sssec:breakdown}

The preceding subsections establish a clean theoretical picture: in a
complete market, the martingale representation theorem provides the
replicating strategy, the Doob--Meyer decomposition separates the hedging
gains from a predictable cost process, and the cost process~$\tilde{C}$
vanishes identically.  The purpose of this final subsection is to
catalogue the specific ways in which this picture breaks down in
practice, thereby motivating the move to incomplete markets in
Section~\ref{subsec:incomplete_markets}.

\paragraph{Dimensional mismatch.}
The predictable representation property~\eqref{eq:PRP} requires that
the market filtration be generated by the same Brownian motions that
drive the traded assets.  When the model includes $m > d$ independent
Brownian motions but only $d$ traded assets, the filtration contains
information generated by the $m - d$ ``extra'' Brownian motions, and
the representation fails.  Concretely, there exist
$\mathbb{Q}$-martingales that cannot be written as stochastic integrals
with respect to the price process~$X$.  These unrepresentable martingales
correspond to risks that no trading strategy can hedge.

In the multi-factor GBM model considered in this thesis, the two-asset
price dynamics are driven by $m = 3$ (or more) independent Brownian
motions.  The volatility matrix $\Sigma$ is $2 \times 3$ and has a
non-trivial null space.  Risks associated with the Brownian motion in this
null space are unhedgeable, and the cost process~$\tilde{C}$ accumulates
whenever these risks affect the option value.

\paragraph{Non-traded risk factors.}
Even when the number of Brownian motions equals the number of traded
assets, the market is incomplete if some risk factors evolve
stochastically but are not themselves tradable.  The canonical example is
the Heston stochastic volatility model \citep{Heston1993}, where the
instantaneous variance of the asset price follows a mean-reverting
diffusion.  The variance process generates a source of randomness that
affects option values --- through the dependence of continuation values on
current and future volatility --- but cannot be hedged because variance is
not a traded instrument.

The mathematical mechanism is as follows.  In the Heston model, the
filtration is generated by two Brownian motions (one driving the asset
price, one driving the variance), but only one instrument is traded.  The
Girsanov theorem requires a two-dimensional market price of risk
$(\gamma^S, \gamma^v)$ but provides only one equation (the drift
condition on the traded asset), leaving $\gamma^v$ --- the variance risk
premium --- as a free parameter.  The drift restriction pins down the
price-risk premium~$\gamma^S$, while the variance-risk
premium~$\gamma^v$ remains underdetermined, producing a family of ELMMs.
Each admissible choice of $\gamma^v$
defines a different element of~$\mathcal{M}(X)$, and the resulting
non-uniqueness is the mathematical signature of the untraded variance
risk.

\paragraph{Other sources of incompleteness.}
Two additional mechanisms are noted for completeness, though they are not
the primary focus of this thesis.  Jump processes
\citep{Merton1976} introduce discontinuities in asset prices that cannot
be replicated by continuous trading, even if the number of Brownian
motions and traded assets match.  Transaction costs
\citep{LelandReplication1985} prevent the continuous rebalancing assumed
by the theory: in the presence of proportional costs, the
self-financing condition is violated at each trade, and perfect
replication is impossible regardless of the model's dimensional structure.

% -----------------------------------------------------------------
\subsubsection{Gap Identification}
\label{sssec:gap_identification}

The classical theory developed in this section handles the case of
complete markets, where the equivalent local martingale measure is unique,
every contingent claim is replicable, and the computational workflow
proceeds in two sequential stages: first solve for the price (via a PDE,
a tree, or a Monte Carlo estimator), then differentiate to obtain the
hedge.  The cost process~$\tilde{C}$ in the optional decomposition
vanishes identically, and the hedging problem reduces to computing a
single sensitivity.

In incomplete markets the situation is fundamentally different.  The set
$\mathcal{M}(X)$ contains more than one element; the Doob--Meyer
decomposition is replaced by the optional decomposition, in which the
cost process~$\tilde{C}$ is strictly positive and merely adapted rather
than predictable; and the relevant value process is no longer the Snell
envelope under a single measure but the minimal dominating wealth process,
defined by a nested supremum over both measures and stopping times.  No
closed-form expressions exist for either the predictable
integrand~$\tilde{\varphi}$ or the cost process~$\tilde{C}$ in this
setting.  Existing numerical methods primarily approximate value functions
under a fixed modelling choice --- often a single measure selected from
$\mathcal{M}(X)$ --- and then differentiate to recover hedges; they do
not directly approximate the pair
$(\tilde{\varphi}, \tilde{C})$ associated with the minimal dominating
wealth process under non-unique martingale measures in high dimension.
Each classical method either targets the price rather than the hedge, or
is restricted to state spaces of dimension one or two, or both.

This thesis addresses the gap by learning $\tilde{\varphi}$
and~$\tilde{C}$ jointly from simulated paths using deep neural networks,
without requiring a prior pricing step and without discretising the state
space on a grid.  The theoretical foundations for the incomplete-market
objects that the network must approximate --- the non-singleton set
$\mathcal{M}(X)$, the minimal dominating wealth process, the optional
decomposition theorem, and the interpretation of~$\tilde{C}$ as the
unhedgeable residual cost --- are developed in the next section.
% To be written

\subsection{Incomplete Markets and Super-Hedging}
\label{subsec:incomplete_markets}

% To be written

\subsection{Optional Decomposition of Supermartingales}
\label{subsec:optional_decomposition}

% To be written

\subsection{Risk Measures and Quantitative Hedging Criteria}
\label{subsec:risk_measures}

% To be written

\subsection{Deep Hedging}
\label{subsec:deep_hedging}

% To be written

\subsection{Market Models}
\label{subsec:market_models}

% To be written

\subsection{Path Signatures as Feature Representations}
\label{subsec:path_signatures}

% To be written

\subsection{Research Gap and Contribution}
\label{subsec:research_gap}

% To be written
