% ============================================================
% Chapter 1: Introduction
% File: chapters/chapter1_introduction.tex
% ============================================================


\clearpage
\section*{Chapter 1}
\section{Introduction}
\label{sec:introduction}
% =========================================================================
% 1.1  HEDGING UNDER NON-UNIQUENESS OF MARTINGALE MEASURES
% =========================================================================
\subsection{Hedging Under Non-Uniqueness of Martingale Measures}
\label{subsec:hedging_motivation}

The modern theory of derivative pricing rests on the principle that, in a
frictionless market with continuous trading, the price of a contingent
claim is determined by an arbitrage argument rather than by the
preferences of individual agents.  In the setting of
\citet{BlackScholes1973} and \citet{Merton1973}, a single risky asset
follows geometric Brownian motion with constant volatility~$\sigma$, and
every European option payoff can be exactly replicated by a dynamically
adjusted portfolio.  The cost of this replicating portfolio is unique: it
equals the discounted expectation of the payoff under the sole probability
measure that makes discounted prices into martingales.

The mathematical foundation is provided by the fundamental theorems of
asset pricing.  The first theorem \citep{DelbaenSchachermayer1994} states
that a market satisfies no free lunch with vanishing risk (NFLVR) if and
only if there exists at least one probability measure
$\mathbb{Q} \sim \mathbb{P}$ under which discounted asset prices are local
martingales.  The set of all such equivalent local martingale measures is
denoted~$\mathcal{M}(X)$.  Financially, each measure
$\mathbb{Q} \in \mathcal{M}(X)$ represents one consistent way of
assigning prices to uncertain future payoffs: it re-weights the likelihood
of different market scenarios so that, on average, discounted asset prices
neither rise nor fall.  The second theorem
\citep{HarrisonPliska1981, HarrisonKreps1979} characterises market
completeness: every contingent claim can be replicated if and only if
$|\mathcal{M}(X)| = 1$.  When a unique measure exists, every derivative
has exactly one arbitrage-free price, and a trading strategy that
perfectly reproduces the payoff can be constructed.

The structural issue that motivates this thesis is the failure of this
uniqueness.  In practice, markets are incomplete for reasons that fall
into two broad categories.  The first is structural incompleteness: a
dimensional mismatch between the number of traded assets and the number of
independent sources of randomness driving the system.  If a market has two
traded assets but three independent Brownian motions, there are not enough
instruments to hedge all sources of risk, regardless of how cleverly the
portfolio is managed.  This form of incompleteness is fixed at model
specification and can, in principle, be resolved by introducing additional
traded instruments.  The second is dynamic incompleteness: the presence of
risk factors that evolve randomly over time but are not themselves
tradable.  Stochastic volatility \citep{Heston1993} is the canonical
example --- the variance of an asset price fluctuates unpredictably, yet
variance itself cannot be bought or sold in the spot market.  This form of
incompleteness persists regardless of the number of traded assets because
the source of risk is intrinsic to the dynamics of the market.  Other
sources of incompleteness include jumps in asset prices
\citep{Merton1976} and transaction costs that prevent continuous
rebalancing \citep{LelandReplication1985}.  In each case, there exist
sources of randomness that cannot be offset by the available instruments,
rendering $|\mathcal{M}(X)| > 1$.  Intuitively, an incomplete market is
one in which there are more sources of risk than there are instruments to
trade them away, so that perfect replication of a general payoff is
impossible.

When the market is incomplete, different measures in~$\mathcal{M}(X)$
assign different prices to the same non-replicable claim.  Rather than a
unique price, one obtains the arbitrage-free price interval
\begin{equation}
\label{eq:price_interval}
\Bigl[\,
  \inf_{\mathbb{Q} \in \mathcal{M}(X)} \mathbb{E}^{\mathbb{Q}}[H],
  \;\;
  \sup_{\mathbb{Q} \in \mathcal{M}(X)} \mathbb{E}^{\mathbb{Q}}[H]
\,\Bigr].
\end{equation}
Any price within this interval is consistent with the absence of
arbitrage; any price outside it creates a free lunch
\citep{DelbaenSchachermayer1994}.  Financially, the interval represents
the range of prices at which neither the buyer nor the seller can extract a
risk-free profit: each measure~$\mathbb{Q}$ yields a different but
legitimate valuation, and the market provides no mechanism to select among
them.  The upper bound
$\pi(H) := \sup_{\mathbb{Q} \in \mathcal{M}(X)}
\mathbb{E}^{\mathbb{Q}}[H]$ has a direct financial interpretation grounded
in martingale duality: it equals the minimal initial capital required to
construct an admissible wealth process that dominates~$H$ almost surely at
maturity.  In other words, it is the cheapest amount of money a seller
must hold today to guarantee that, by trading in the available assets, the
portfolio will cover the payoff regardless of which scenario materialises.
This is the super-hedging price.  In a complete market, the interval
degenerates to a single point because only one pricing measure exists;
computationally, this means a single expectation suffices.  In an
incomplete market, computing the upper bound requires an optimisation over
the entire set~$\mathcal{M}(X)$, which is generally infinite-dimensional.

American options compound this difficulty through optimal stopping.  The
holder may exercise at any stopping time $\tau \le T$, and the seller's
hedge must remain adequate regardless of when exercise occurs.  The hedging
problem is therefore coupled to an optimal stopping problem, and the
minimal dominating price acquires a minimax structure:
\begin{equation}
\label{eq:american_price_intro}
\pi(\xi)
\;:=\;
\sup_{\mathbb{Q} \in \mathcal{M}(X)}
\;\sup_{\tau \in \mathcal{T}_{[0,T]}}
\;\mathbb{E}^{\mathbb{Q}}[\xi_\tau],
\end{equation}
where $\xi := (\xi_t)_{t \in [0,T]}$ is the discounted reward process and
$\mathcal{T}_{[0,T]}$ is the set of stopping times valued in~$[0,T]$.
The nested supremum creates a two-player adversarial structure: the outer
supremum searches over all consistent pricing models (choosing the
worst-case measure), while the inner supremum searches over all possible
exercise times (choosing the worst-case exercise for the seller).  The
seller must be prepared for the worst combination of both.  Shifting the
measure~$\mathbb{Q}$ alters the expected continuation value of the option
at each point in time, which in turn shifts the optimal stopping boundary:
under a measure that inflates the probability of adverse price movements,
early exercise becomes more attractive, increasing the seller's exposure.
In complete markets, the outer supremum is trivial (there is only one
measure) and the problem reduces to the classical Snell envelope
\citep{Bensoussan1984, Karatzas1988}, a single optimal stopping problem.
In incomplete markets, both suprema are non-trivial: the seller faces
adversarial choice over both the pricing measure and the exercise policy,
making the problem substantially harder.  Computationally, this means that
any numerical method must handle the interaction of two coupled
optimisation layers, which rules out straightforward dynamic programming
unless the model has special structure.

The central object of this thesis is the minimal dominating wealth process
under non-unique martingale measures: the adapted process that tracks this
value dynamically through time, together with the trading strategy that
generates it.  The following section formulates the problem precisely and
identifies the mathematical gap that the thesis addresses.


% =========================================================================
% 1.2  MATHEMATICAL FORMULATION
% =========================================================================
\subsection{Mathematical Formulation of the Problem}
\label{subsec:math_formulation}

Let $(\Omega, \mathcal{F},
\mathbb{F} := (\mathcal{F}_t)_{t \ge 0}, \mathbb{P})$ be a filtered
probability space satisfying the usual conditions of right-continuity and
completeness.  The filtration $\mathbb{F}$ represents the information
available to the hedger at each point in time: $\mathcal{F}_t$ contains
everything that has been observed up to and including time~$t$.  The space
supports $d = 2$ traded risky assets with discounted price process
$X := (X^1, X^2)$, assumed to be a locally bounded, c\`{a}dl\`{a}g
(right-continuous with left limits) semimartingale.  The undiscounted price
of asset~$i$ is denoted~$S^i_t$, and the discounted price is
$X^i_t := e^{-rt} S^i_t$ where~$r$ is the risk-free rate.  Let
$\mathcal{M}(X)$ denote the set of all probability measures
$\mathbb{Q} \sim \mathbb{P}$ under which~$X$ is a local martingale.  The
market is assumed to be incomplete: $|\mathcal{M}(X)| > 1$, reflecting
the presence of $m > d = 2$ independent sources of randomness.  In the
models considered in this thesis, there are two traded assets but three or
more independent Brownian motions driving the system, so that perfect
hedging is impossible.

A trading strategy is a predictable, $X$-integrable process
$\varphi := (\varphi^1, \varphi^2)$ representing the number of units of
each asset held at each time.  Predictability means that the portfolio
position at time~$t$ is determined using only information available
strictly before~$t$: the hedger must decide how much of each asset to hold
before observing the next price movement.  Given initial capital $v \ge 0$
and strategy~$\varphi$, the wealth process is
\begin{equation}
\label{eq:wealth_process}
V_t
\;:=\; v \;+\; \int_0^t \varphi_s \, dX_s \;-\; C_t,
\qquad t \ge 0,
\end{equation}
where $\int_0^t \varphi_s \, dX_s$ represents the cumulative gains from
trade and $C := (C_t)_{t \ge 0}$ is an adapted, non-decreasing process
with $C_0 = 0$.  Financially, the integral $\int_0^t \varphi_s \, dX_s$
is the total profit or loss from holding the portfolio: it accumulates the
gains and losses from each small price movement, weighted by the number of
shares held.  The process~$C$ records the cumulative cost that arises
because the market does not offer enough instruments to track the
obligation perfectly.  When $C \equiv 0$, the strategy is self-financing:
changes in wealth are driven solely by changes in asset prices, and no
external cash injection is needed.  When $C$ is strictly increasing, the
hedger must absorb a running cost, reflecting the irreducible price of
operating in an incomplete market.  The strategy is admissible if
$V_t \ge 0$ for all $t \ge 0$ \citep{Kramkov1996}; this rules out
strategies that borrow indefinitely.

The American put option with strike~$K$ and maturity~$T$ on asset~$i$
generates a discounted reward process
\begin{equation}
\label{eq:american_reward}
\xi_t \;:=\; e^{-rt}\max(K - S^i_t,\, 0),
\qquad t \in [0, T].
\end{equation}
This is the payoff the holder would receive if exercising at time~$t$,
discounted back to time zero.  An admissible strategy with
$V_t \ge \xi_t$ for all $t \in [0,T]$ is a dominating strategy
for~$\xi$: the portfolio value always exceeds the option obligation, so
the seller can meet the claim no matter when the holder exercises.  The
minimal dominating value process is
\begin{equation}
\label{eq:value_process}
\tilde{V}_t
\;:=\;
\esssup_{\mathbb{Q} \in \mathcal{M}(X)}
\;\esssup_{\tau \in \mathcal{T}_{[t,T]}}
\;\mathbb{E}^{\mathbb{Q}}[\xi_\tau \mid \mathcal{F}_t].
\end{equation}
The essential supremum ($\esssup$) is the measure-theoretic analogue of a
pointwise maximum.  It is needed because the supremum over an uncountable
family of random variables need not itself be measurable; the essential
supremum is the smallest random variable that dominates every member of
the family almost surely.  Intuitively, $\tilde{V}_t$ is the smallest
amount of wealth that the seller must hold at time~$t$ to guarantee
coverage of the American option from that point onward, accounting for the
worst-case pricing model and the worst-case exercise time.  At time zero,
$\tilde{V}_0$ reduces to the minimal dominating price~$\pi(\xi)$.

Under the assumption that $\tilde{V}$ is a $\mathbb{Q}$-supermartingale
for every $\mathbb{Q} \in \mathcal{M}(X)$ a condition satisfied by
the American put payoff~\eqref{eq:american_reward} under the market models
considered in this thesis, since the discounted reward process is
non-negative and bounded above by~$K$, ensuring uniform integrability
the optional decomposition theorem \citep{Kramkov1996} guarantees
that~$\tilde{V}$ admits the representation
\begin{equation}
\label{eq:optional_decomp}
\tilde{V}_t
\;=\; \tilde{V}_0
\;+\; \int_0^t \tilde{\varphi}_s \, dX_s
\;-\; \tilde{C}_t,
\qquad t \ge 0,
\end{equation}
where $\tilde{\varphi}$ is a predictable, $X$-integrable process and
$\tilde{C}$ is an adapted, non-decreasing, c\`{a}dl\`{a}g process with
$\tilde{C}_0 = 0$.  Note that~\eqref{eq:optional_decomp} is not a
definition: $\tilde{V}_t$ is defined by the essential
supremum~\eqref{eq:value_process}, and the
decomposition~\eqref{eq:optional_decomp} is a representation result that
follows from the theorem.  This theorem is the cornerstone of the thesis
and merits careful interpretation.  It states three things.  First, the
minimal dominating value can be decomposed into a starting capital
$\tilde{V}_0$, a gains-from-trade component
$\int_0^t \tilde{\varphi}_s \, dX_s$, and a cost
component~$\tilde{C}_t$.  Second, the trading
strategy~$\tilde{\varphi}$ is predictable, meaning it can be determined
before observing the current price movement.  Third, the cost
process~$\tilde{C}$ is only adapted: its value at time~$t$ may depend on
information revealed at~$t$ itself, not merely on information available
before~$t$.  This is why the decomposition is called optional rather than
predictable, in contrast to the classical Doob--Meyer decomposition in
which the increasing process is predictable \citep{KaratzasShreve1991}.

Financially, the decomposition separates two fundamentally different
components of the hedging problem.  The integral
$\int_0^t \tilde{\varphi}_s \, dX_s$ captures the part of the value
process that can be generated by trading: it is what the hedger earns
through skilful portfolio management.  The process~$\tilde{C}_t$ captures
the part that cannot be generated by trading: it is the cumulative cost of
operating in an incomplete market, the price paid for the fact that the
available instruments do not span all sources of risk.  In the Heston
model, for example, $\tilde{C}$ accumulates when the variance process
moves in a direction that the hedger cannot offset because variance is not
directly tradable.  In a complete market, $\tilde{C} \equiv 0$ and the
decomposition reduces to self-financing replication: the entire value
process is generated by trading alone.  In an incomplete market,
$\tilde{C}$ is strictly increasing and measures the irreducible cost of
the gap between available instruments and the risks to be managed.

The distinction between predictable and adapted is central to this thesis
and has a direct computational consequence.  The
integrand~$\tilde{\varphi}$ is predictable: the hedger's portfolio
position at time~$t$ depends only on information available strictly
before~$t$.  The cost process~$\tilde{C}$ is adapted but generally not
predictable: its value at time~$t$ may depend on information revealed
at~$t$ itself, such as a sudden volatility shock or an unexpected
correlation change.  A neural network that maps the current market state to
a portfolio position at each rebalancing date is, by construction, an
estimator of the predictable integrand~$\tilde{\varphi}$: it produces a
trading decision based on information available at the time the decision is
made.  The residual hedging error the gap between the trained portfolio
value and the option obligation captures the component of the value
process that the predictable strategy cannot reach.  This residual serves
as the empirical proxy for the adapted cost process~$\tilde{C}$.

It should be noted that the predictable/adapted distinction is exact in
continuous time.  In the discrete-time implementation used in this thesis,
the distinction becomes a modelling choice at each rebalancing date~$t_k$:
the network receives the state observed at~$t_k$ and produces a position
that is held until~$t_{k+1}$.  The convention adopted throughout is that
the network's input at step~$k$ consists of information available at the
start of the interval~$[t_k, t_{k+1})$, so that the resulting strategy is
predictable with respect to the discrete filtration.  

The theorem is existential.  It guarantees the existence
of~$\tilde{\varphi}$ and~$\tilde{C}$ but provides no constructive
algorithm for computing either.  It tells us that an optimal predictable
strategy exists and that the remaining cost is adapted, but it does not
tell us what the strategy looks like in any specific model.  This gap
between the existence guarantee and the need for a computable solution
defines the problem addressed in this thesis: to
approximate~$\tilde{\varphi}$ using established neural network methods, and
to interpret the residual hedging error as an empirical estimate
of~$\tilde{C}$.


% =========================================================================
% 1.3  COMPUTATIONAL CHALLENGE AND RESEARCH GAP
% =========================================================================
\subsection{Computational Challenge and Research Gap}
\label{subsec:computational_challenge}

Computing the optional decomposition in concrete settings requires solving
high-dimensional, path-dependent optimisation problems for which no
closed-form solutions exist.  Classical PDE methods apply when the price
process is Markovian and low-dimensional, but scale poorly with the number
of state variables: in the two-asset Heston model considered here, the
state space is at least four-dimensional (two asset prices and two
variance processes), and grid-based methods become impractical
\citep{Glasserman2003}.  Moreover, PDE-based value functions depend on the
choice of pricing measure, and no single PDE characterises the minimal
dominating value process without additional structural information
about~$\mathcal{M}(X)$.  Regression-based Monte Carlo methods, notably the
Longstaff--Schwartz algorithm \citep{LongstaffSchwartz2001}, address
optimal stopping but do not directly yield hedging strategies and do not
account for model ambiguity in incomplete markets.

Deep learning methods offer alternatives that bypass these limitations.
Two fundamentally different approaches are relevant, distinguished by
whether they solve the problem backward or forward in time.

The deep BSDE solver \citep{HanJentzenE2018} approximates the backward
stochastic differential equation (BSDE) that characterises the value
process.  Given a terminal condition (the option payoff), it propagates the
solution back through the time grid, parameterising the initial
value~$Y_0$ and the control process~$Z_{t_k}$ at each discrete step with
feedforward sub-networks.  The structural analogy to the optional
decomposition is informative: the process~$Y$ plays the role of the
wealth~$V$, the control~$Z$ plays the role of the predictable
integrand~$\tilde{\varphi}$, and the driver of the backward equation
encodes a cost structure analogous to~$\tilde{C}$.  However, this
correspondence should be understood with care.  The deep BSDE solver
operates under a single reference measure and solves a specific non-linear
backward equation, whereas the optional decomposition characterises the
value process simultaneously across the entire
set~$\mathcal{M}(X)$ via an essential supremum over all equivalent local
martingale measures.  The alignment between the BSDE solution and the
optional decomposition is therefore structural rather than exact: the
network learns an object that is analogous to~$\tilde{\varphi}$ within the
reference model, but the extent to which this approximation captures the
full supremum over~$\mathcal{M}(X)$ is an empirical question rather than a
guaranteed mathematical identity.

By contrast, the deep hedging framework
\citep{BuehlerGononTeichmannWood2019} operates forward in time: it
parameterises the trading strategy directly as a neural network and trains
it end-to-end by minimising a risk functional applied to the distribution
of hedging errors over simulated paths.  Rather than solving a backward
equation, it searches directly over strategies to find one whose terminal
hedging error is small.  The network learns~$\tilde{\varphi}$ indirectly:
the loss function penalises the terminal hedging error, and
backpropagation through the trading sequence adjusts the network weights so
that the resulting strategy approximates the optimal predictable integrand.
Within this forward framework, two architectural choices are considered: a
GRU recurrent network \citep{Cho2014} that maintains an evolving hidden
state updated at each observation, and a feedforward network with path
signature features \citep{Horvath2021} that constructs a static,
fixed-dimensional mathematical summary of the path history.  A detailed
review of these methods and their extensions is deferred to the following Chapters.

The distinction between direct and indirect learning
of~$\tilde{\varphi}$ has consequences for how each architecture captures
the predictable and adapted components of the optional decomposition.  The
deep BSDE solver directly parameterises the control~$Z$ that corresponds
to~$\tilde{\varphi}$ within its reference model.  The forward
architectures optimise a risk functional over the terminal hedging error,
and the approximation of~$\tilde{\varphi}$ emerges as a byproduct of
minimising that error.  In both cases, whatever the network cannot learn
ends up in the residual, which reflects the cost
process~$\tilde{C}$.  Whether the GRU's recurrent memory, which updates
its hidden state with information revealed at each time step, offers an
advantage in capturing the adapted nature of~$\tilde{C}$ relative to the
static signature encoding is an empirical question that the thesis
investigates.

The three architectures considered in this thesis : deep BSDE, GRU, and
FNN with path signatures originate from distinct research communities.
Deep BSDE solvers emerged from the numerical PDE and backward stochastic
differential equation literature \citep{HanJentzenE2018, PardouxPeng1990},
where the primary applications have been high-dimensional semilinear PDEs
and European-style claims.  Deep hedging originated in quantitative finance
and risk management \citep{BuehlerGononTeichmannWood2019}, with a focus on
transaction costs, market frictions, and regulatory constraints for
European portfolios.  Path signature methods entered computational finance
from rough path theory \citep{Horvath2021, LyonsCaruanaLevy2007}, initially as
feature extractors for volatility modelling rather than as hedging
architectures.  Each method has been developed, tested, and validated
within its own community, against its own baselines, and under its own
experimental conventions.  

As a result, reported performance differences across studies may reflect differences in market models, loss functions, hyperparameter budgets, or evaluation protocols rather than genuine
architectural capabilities.  Furthermore, American options under incomplete
markets compound two difficulties, optimal stopping and model ambiguity
, that most existing studies avoid by restricting attention to European
claims or to complete market settings.  The combination of early exercise,
non-unique martingale measures, and the optional decomposition perspective
has therefore not served as a test bed for any of these methods, let alone
for a controlled comparison among them.

\subsubsection*{Rationale for the Architectural Comparison}
Comparing all three under identical conditions is important for three
reasons.  First, the architectures encode fundamentally different inductive
biases: the deep BSDE solver imposes backward equation structure and
learns~$\tilde{\varphi}$ through the control~$Z$; the GRU learns a
forward strategy with recurrent memory that may capture path-dependent
features; the FNN with path signatures compresses the path history into a
static algebraic summary before mapping it to a decision.  These are not
minor implementation variants; they represent distinct hypotheses about
which representation of market information leads to the best approximation
of the predictable integrand.  Second, without a unified experimental
design, it is impossible to determine whether observed differences are due
to the architecture itself or to confounding factors such as the choice of
loss function, the number of training paths, or the hyperparameter search
procedure.  Third, the optional decomposition provides a shared
theoretical benchmark: by interpreting each architecture's output as an
approximation of~$\tilde{\varphi}$ and its residual as an estimate
of~$\tilde{C}$, the comparison becomes scientifically meaningful rather
than a purely engineering exercise.  Without this common interpretive
framework, comparing architectures would reduce to reporting loss values
with no structural insight into what each method has and has not captured.

\subsubsection*{Research gaps}
Against this background, three specific gaps remain in the existing
computational literature.  First, no systematic comparison of these
architectures has been conducted for American options under controlled
incomplete-market models, where the neural network output is explicitly
interpreted as the estimator of~$\tilde{\varphi}$ and the residual error
as the empirical proxy for~$\tilde{C}$.  Deep hedging studies have
primarily addressed European claims under transaction costs and market
frictions; applications to American options under model ambiguity within
the optional decomposition framework are absent.  Second, the distinct
contributions of structural incompleteness and dynamic incompleteness
(as defined in Section~\ref{subsec:hedging_motivation}) to the residual
hedging cost have not been disentangled empirically.  Understanding which
source dominates the hedging cost informs both model selection and
instrument design.  Third, the residual error of trained hedging models has
not been validated against the theoretical prediction that
$\tilde{C} \equiv 0$ in the complete-market limit and $\tilde{C} > 0$
otherwise.


% =========================================================================
% 1.4  EVALUATION FRAMEWORK AND LOSS FUNCTION
% =========================================================================
\subsection{Evaluation Framework and Loss Function}
\label{subsec:loss_function}

\subsubsection{Loss function and actuarial rationale}
All three architectures are trained using a common asymmetric loss
function.  To state it precisely, denote by~$\varphi_\theta$ a
neural-network-parameterised trading strategy with parameter
vector~$\theta$.  Given initial capital~$v$, the parameterised wealth
process is
\begin{equation}
\label{eq:param_wealth}
V_t^\theta
\;:=\; v \;+\; \int_0^t \varphi_\theta(s) \, dX_s,
\qquad t \ge 0.
\end{equation}
The process~$V^\theta$ is the neural network's approximation of the
theoretical minimal dominating wealth~$\tilde{V}$: both are wealth
processes generated by a predictable trading strategy applied to the same
underlying assets, but~$\tilde{V}$ is the exact optimum guaranteed by the
optional decomposition theorem while~$V^\theta$ is the learned
approximation parameterised by network weights.  The loss function is

\begin{equation}
\label{eq:loss}
\mathcal{L}(\theta)
\;:=\; \lambda_s \, \mathbb{E}\bigl[\max(\xi_\tau - V_T^\theta,\, 0)\bigr]
\;+\; \lambda_o \, \mathbb{E}\bigl[\max(V_T^\theta - \xi_\tau,\, 0)\bigr]
\;+\; \lambda_c \, \operatorname{CVaR}_\alpha\!\bigl(\max(\xi_\tau
      - V_T^\theta,\, 0)\bigr),
\end{equation}
where $\operatorname{CVaR}_\alpha$ denotes conditional value-at-risk at
level~$\alpha$ \citep{RockafellarUryasev2000}.  

The three terms encode distinct financial objectives.The first term, weighted by~$\lambda_s$, penalises shortfall: scenarios in which the portfolio fails to cover the option obligation.  Financially, this is a solvency constraint.  A shortfall $\xi_\tau > V_T^\theta$ means the hedger cannot meet the liability, which constitutes a default event. The second term, weighted by~$\lambda_o$, penalises over-hedging: scenarios in which the portfolio exceeds the obligation.  Financially,
this is a capital efficiency constraint.  A surplus $V_T^\theta > \xi_\tau$ means the hedger has immobilised more capital than necessary, incurring an opportunity cost but without threatening solvency. The asymmetry $\lambda_s \gg \lambda_o$ \footnote{In practice, the
penalty on shortfall is chosen significantly larger than the penalty
on surplus to reflect regulatory capital requirements, funding
constraints, and the severe reputational and legal consequences of
default.  Over-hedging merely entails an opportunity cost, whereas
shortfall corresponds to a solvency event.} reflects the practical reality that default is far more costly than capital inefficiency.

The third term targets the tail of the shortfall distribution using CVaR
rather than standard Value-at-Risk.  Standard VaR reports only the
threshold of a given quantile: it answers ``what is the worst loss at the
95th percentile?''\ but is blind to how bad losses can be beyond that
threshold.  CVaR averages over the entire tail beyond the quantile,
penalising the severity of the worst outcomes rather than merely their
frequency \citep{Artzner1999}.  For a hedger concerned with catastrophic
shortfall scenarios, CVaR is the appropriate measure because it is
sensitive to the shape of the tail, not just its location.

This loss does not enforce the strict domination $V_t \ge \xi_t$ for
all~$t$ that would characterise a full super-replicating strategy.
Instead, it encodes a soft preference for strategies whose shortfall is
small and whose tail risk is controlled.  From the perspective of the
optional decomposition, the mean residual hedging error provides an
empirical estimate of the expected value of~$\tilde{C}$, while the
shortfall distribution characterises the pathwise behaviour
of~$\tilde{C}$ across scenarios.  Calibration of the penalty parameters
$\lambda_s$, $\lambda_o$, and $\lambda_c$ is described in the following Chapters.

\subsubsection{Path generation and robustness.}

Training requires large numbers of simulated price paths. Alongside standard parametric simulation from the GBM and Heston models, this thesis employs a variational autoencoder (VAE) trained on historical multi-asset equity data to generate synthetic trajectories. The VAE serves as a stress test for model misspecification because it captures empirical distributional features such as volatility clustering, fat tails, and cross-asset dependence structures that the parametric models may fail to reproduce.

By evaluating hedging performance on both parametric and VAE-generated paths, one can assess the extent to which conclusions depend on the assumed data-generating process. If a hedging strategy performs well on parametric paths but poorly on VAE-generated paths, this indicates that the strategy has overfit to model-specific features rather than learning a robust approximation of~$\tilde{\varphi}$.


% =========================================================================
% 1.5  RESEARCH QUESTIONS
% =========================================================================
\subsection{Research Questions}
\label{subsec:research_questions}

The thesis addresses the following empirical research questions.

\begin{enumerate}
\item \underline{Approximation of the minimal dominating wealth process. }
  Can neural-network-based trading strategies, interpreted as estimators
  of the predictable integrand~$\tilde{\varphi}$, approximate the minimal
  dominating wealth process implied by martingale duality in multi-asset
  incomplete markets? How closely do the deep BSDE solver, GRU network,
  and FNN with path signature features track the value process, and how
  do their residual errors compare under identical conditions?

\item \underline{Structural versus dynamic incompleteness. }
  How does the source of market incompleteness
  (as defined in Section~\ref{subsec:hedging_motivation}) affect the
  magnitude and distribution of the residual hedging error? The
  multi-factor GBM model introduces structural incompleteness through a
  dimensional mismatch between traded assets and Brownian drivers, while
  the Heston model introduces dynamic incompleteness through the random
  evolution of untraded variance processes. Comparing the two isolates
  the contribution of each mechanism to the total hedging cost.

\item \underline{Recurrent memory and adapted cost components. }
  The cost process~$\tilde{C}$ in the optional decomposition is adapted
  but generally not predictable: its value at time~$t$ may depend on
  information revealed at~$t$ itself. The GRU architecture maintains an
  evolving hidden state that is updated with each new observation. The
  FNN with path signatures constructs a static, fixed-dimensional summary
  of the path history computed from data available strictly before the
  decision point. Is the GRU's recurrent memory better suited to
  approximating the non-predictable, adapted component of the cost
  process~$\tilde{C}$ than the static signature encoding?

\item \underline{Backward versus forward deep learning approaches. }
  The deep BSDE solver approximates the hedging problem through a
  backward equation whose control variable corresponds
  to~$\tilde{\varphi}$ within its reference model. The GRU and FNN-Sig
  architectures parameterise the strategy directly and optimise a forward
  risk functional, learning~$\tilde{\varphi}$ indirectly through the
  gradient of the loss. How do these approaches compare in mean hedging
  error, tail risk (measured by CVaR), and computational cost?

\item \underline{Robustness to model misspecification.} 
  How do hedging models trained on VAE-generated paths, which reflect
  empirical distributional features learned from historical data, compare
  with those trained on parametric (GBM, Heston) simulations? Does
  data-driven path generation, serving as a stress test for model
  assumptions, improve out-of-sample hedging performance?
\end{enumerate}


Each question is empirical. The answers depend on measurable quantities such as the mean hedging error, shortfall probability, CVaR, and the distribution of the residual cost, all of which are estimated from simulated experiments with controlled statistical uncertainty.



% =========================================================================
% 1.6  CONTRIBUTIONS
% =========================================================================
\subsection{Contributions}
\label{subsec:contributions}

The contributions of this thesis are computational and empirical.  All
mathematical results, pricing models, representation theorems, and
numerical solvers are drawn from the established literature.

\begin{enumerate}
\item \underline{Empirical proxy for the cost of incompleteness.}
  The neural network output at each rebalancing date is interpreted as
  the estimator of the predictable integrand~$\tilde{\varphi}$, and the
  residual hedging error is interpreted as the empirical counterpart of
  the adapted cost process~$\tilde{C}$ in the optional decomposition.
  This is the primary contribution of the thesis.  It provides empirical
  evidence consistent with the theoretical prediction that~$\tilde{C}$ is
  strictly positive in incomplete markets and approximately zero in the
  complete-market limit, thereby validating the optional decomposition as
  a practical diagnostic framework for hedging under model ambiguity.

\item \underline{Comparative evaluation under identical conditions.}
  A systematic empirical comparison of three neural-network-based hedging
  architectures, namely the deep BSDE solver, a GRU recurrent network, and a feedforward network with path signature features, applied to American
  put options under the same market models, loss function, and evaluation
  protocol. Because these architectures originate from distinct research
  communities, namely numerical PDEs, quantitative risk management, and rough path theory, existing studies evaluate them against different baselines and under different experimental conventions. The optional decomposition provides a shared theoretical benchmark that makes the comparison scientifically meaningful rather than a purely engineering exercise. A two-stage experimental design consisting of Bayesian hyperparameter optimisation \citep{Snoek2012} followed by retraining across multiple random seeds \citep{Henderson2018, Bouthillier2021} ensures that reported performance reflects model capability rather than fortunate initialisation.

\item \underline{Two-model study isolating sources of incompleteness.}
  By comparing a multi-driver GBM model (structural incompleteness
  through dimensional mismatch) with a two-asset Heston model (dynamic
  incompleteness through stochastic volatility), the thesis disentangles
  the contributions of each mechanism to the residual hedging error and
  provides empirical evidence on how the source of incompleteness, not
  merely its presence, affects hedging difficulty.

\item \underline{Tail-risk-based evaluation with actuarial rationale.}
  Performance is evaluated using the asymmetric loss function with a CVaR
  component, providing a risk-sensitive assessment that characterises the
  tail of the shortfall distribution.  The solvency-oriented framing of
  the shortfall penalty and the capital-efficiency interpretation of the
  over-hedging penalty align the evaluation with actuarial risk management
  principles.

\item \underline{VAE-based path generation as a stress test.}
  A variational autoencoder trained on historical multi-asset equity data
  provides a data-driven complement to parametric simulation.  By
  capturing volatility clustering, fat tails, and cross-asset dependence
  that parametric models may not reproduce, the VAE-generated paths serve
  as a stress test for model misspecification, enabling assessment of the
  robustness of hedging performance across data-generating assumptions.
\end{enumerate}

The thesis uses established methods as computational tools and evaluates
them on a problem, namely multi-asset American options under controlled
incomplete-market models, that has not been studied in this combination
of models, instruments, and architectures.  The optional decomposition
provides the shared theoretical benchmark that gives the comparison
structural content. Without it, cross-architecture evaluation would reduce
to reporting loss values with no insight into what each method has and has
not captured about the hedgeable and unhedgeable components of the problem.

% =========================================================================
% 1.6  STRUCTURE OF THE THESIS
% =========================================================================
\subsection{Structure of the Thesis}
\label{subsec:thesis_outline}

